---
title: "STA 9750 Projects"
execute:
  echo: false
  message: false
  warning: false
---

## About me

I’m **Apu Datta**, a Master’s student in Business Analytics at *Baruch College, CUNY*. I hold a **BBA in Accounting** and an **MBA in Finance** from *Independent University, Bangladesh*, both recognized as U.S. equivalent degrees.  

With over a decade of experience in finance, ERP systems (**SAP**), and analytics, I have served as Assistant General Manager (Financial Planning Analyst) at **Epic Group, Bangladesh**. I also led the **SAP S/4HANA** implementation at **Deshbandhu Group** and contributed to corporate finance at **VIYELLATEX Group, Bangladesh**.

## Baruch College Map

```{r}
if (!require("leaflet")) {
  options(repos = c(CRAN = "https://cloud.r-project.org"))
  install.packages("leaflet")
  stopifnot(require("leaflet"))
}

# Baruch College (NYC)
baruch_longitude <- -73.98333
baruch_latitude  <- 40.75028

leaflet() |>
  addTiles() |>
  setView(baruch_longitude, baruch_latitude, zoom = 17) |>
  addPopups(baruch_longitude, baruch_latitude,
            "I am a Master's student at <b>Baruch College</b>!")

```

## Independent University, Bangladesh (Dhaka)

```{r}

iub_longitude <- 90.4275
iub_latitude  <- 23.8155

leaflet() |>
  addTiles() |>
  setView(iub_longitude, iub_latitude, zoom = 16) |>
  addPopups(iub_longitude, iub_latitude,
            "I earned my <b>BBA</b> and <b>MBA in Finance</b> at <b>Independent University, Bangladesh</b>.")

```

## More About Me

- **Professional Interests**: Forecasting, Financial Analytics, ERP systems (SAP), Machine Learning, Data visualization
- **Tools**: Excel, SAP, Python, SQL, Git/GitHub, R
- **Contact**: apu.datta@baruchmail.cuny.edu
- **LinkedIn**: [linkedin.com/in/apu-datta-491b8b84](https://www.linkedin.com/in/apu-datta-491b8b84)  
- **GitHub**: [github.com/dattaBus-anls](https://github.com/dattaBus-anls)
- **Resume**: [Open my resume](Resume.pdf)

```{r}
#| include: false
1+1
```

Last Updated: `r format(Sys.time(), "%A %b %d, %Y at %I:%M %p")`


